---
title: "tidyverse and ggplot"
author: "Afif Amirah"
date: "2/9/2022"
output: 
  html_document :
    number_sections : yes
    theme : cerulean
    highlight : kate
    toc : yes
    toc_float :
      collapsed: true
      smooth_scroll: true
  word_document :
    toc : yes
  pdf_document :
    toc : yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Import Data

We will be looking at a data frame called 'democracy_data' and how to go about wrangling the data set into digestable information. Data source website <https://moderndive.com/data/dem_score.csv>.

First step, we are going to properly filter the data to focus on information we want to look at, say we only want to look at democracy scores for country = "Guatemala" :

```{r echo=TRUE, message=FALSE, warning=FALSE}
democracy_data <- read_csv("https://moderndive.com/data/dem_score.csv",show_col_types = FALSE)
guatemala_demo <- democracy_data %>% 
  filter(country == "Guatemala")
as.data.frame(guatemala_demo)

```
## Tidy Data

Next, we want to be able to convert the above data frame into 'tidy' format, which means each variable forms a column, each observation forms a row and each observational unit forms a table. Let's see how we can do this transformation :

```{r echo=TRUE, message=FALSE, warning=FALSE}
guat_tidy <- gather(data = guatemala_demo, 
                    key = year,
                    value = democracy_score,
                    - country) 
guat_tidy
```

Now we can see what the 'gather()' function does to the initial data frame is that it converts the years data from multiple columns into one column and corresponds to it is the democracy scores. This 'tidy' format now is nicer to look at.


## Plots

We can now create a decent graph to represent the data frame above :

```{r echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data=guat_tidy, mapping = aes (x=parse_number(year),
                                            y = democracy_score)) +
  geom_line() + labs(x="year",y="democracy_score", title = "Democracy Scores in Guatemala")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
